% Draft results section for the Comp Phys final paper.
% This snippet is intended to be pasted into an Overleaf manuscript.

\section{Results}
\label{sec:results}

\textcolor{red}{Discuss the difficulties and failures of the Unlabeled Shift Metrics -> Accuracy Shift metrics. Also discuss how we weren't able to get runs using Integrated Gradients or SmoothGrad as in-training regularizers because they took far too long when used that way...but we still implemented methods using them. Also discuss all the sanity checks. Discuss a lot of stuff...but I just wanted to flag this stuff as being important to discuss}

\subsection{Reimplementation Protocol and Evaluation Scope}
We reimplemented the preliminary JetClass studies from scratch and evaluated the pipeline across five independent seeds ($\{41,52,63,74,85\}$), with fixed data splits and model/training settings per run. For each seed, we measured: (i) clean offline classification performance, (ii) corruption-induced performance degradation and unlabeled shift metrics, and (iii) interpretability-based masking effectiveness. We then aggregated all metrics across seeds using mean, standard deviation, and min/max.

All five seeded runs completed successfully (5/5 with no missing outputs). This section reports the aggregated results from \texttt{prelim\_reimpl\_cluster\_aggregate\_5seeds}.

\subsection{Clean Baseline Performance}
Table~\ref{tab:clean-aggregate} shows stable clean performance across seeds. The macro one-vs-rest AUC is high ($0.9651\pm0.0019$), with moderate seed-to-seed spread in top-1 accuracy ($0.7459\pm0.0073$). Signal-vs-background and target-vs-background operating-point metrics are similarly consistent.

\begin{table}[t]
\centering
\caption{Clean performance aggregated over 5 seeds.}
\label{tab:clean-aggregate}
\begin{tabular}{lccc}
\toprule
Metric & Mean $\pm$ Std & Min & Max \\
\midrule
Accuracy & $0.7459 \pm 0.0073$ & $0.7369$ & $0.7551$ \\
Macro AUC (OvR) & $0.9651 \pm 0.0019$ & $0.9628$ & $0.9674$ \\
Signal-vs-BG AUC & $0.9573 \pm 0.0015$ & $0.9552$ & $0.9587$ \\
Signal-vs-BG FPR@50\% TPR & $0.00785 \pm 0.00050$ & $0.00727$ & $0.00840$ \\
Target-vs-BG-ratio AUC & $0.9910 \pm 0.0023$ & $0.9877$ & $0.9934$ \\
Target-vs-BG-ratio FPR@50\% TPR & $0.000747 \pm 0.000328$ & $0.000333$ & $0.001067$ \\
Best epoch & $28.8 \pm 1.3$ & $27$ & $30$ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Unlabeled Shift-Metric Calibration and Ranking}
We evaluated five unlabeled shift metrics against corruption-induced $\Delta$AUC. Table~\ref{tab:shift-calib} shows that \texttt{prob\_l1\_drift}, \texttt{top1\_flip\_rate}, and \texttt{class\_js\_divergence} are the strongest rank-correlated indicators of degradation (mean Spearman $\rho\approx0.957$). By contrast, \texttt{confidence\_drop} is weaker and more variable.

Across all five seeds, \texttt{prob\_l1\_drift} ranked first in absolute Spearman correlation with $\Delta$AUC (5/5 seeds), making it the most stable top-ranked unlabeled shift indicator in this reimplementation.

\begin{table}[t]
\centering
\caption{Calibration of unlabeled shift metrics against corruption-induced $\Delta$AUC (5-seed aggregate).}
\label{tab:shift-calib}
\begin{tabular}{lcc}
\toprule
Shift metric & Spearman($\Delta$AUC) mean $\pm$ std & Pearson($\Delta$AUC) mean $\pm$ std \\
\midrule
\texttt{prob\_l1\_drift} & $0.9567 \pm 0.0149$ & $0.9851 \pm 0.0009$ \\
\texttt{top1\_flip\_rate} & $0.9567 \pm 0.0149$ & $0.9849 \pm 0.0010$ \\
\texttt{class\_js\_divergence} & $0.9567 \pm 0.0149$ & $0.9296 \pm 0.0072$ \\
\texttt{entropy\_shift} & $0.8467 \pm 0.0701$ & $0.6443 \pm 0.1767$ \\
\texttt{confidence\_drop} & $0.7600 \pm 0.1512$ & $0.4313 \pm 0.2593$ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Interpretability Method Effectiveness}
We compared targeted masking (mask highest-attribution constituents) to random masking at matched mask fractions. The masking parity check confirms fairness of the intervention size: targeted and random remove the same average number of constituents per jet at each fraction.

Table~\ref{tab:interp-effect} shows that all three methods produce substantial targeted-vs-random gaps, with Integrated Gradients (IG) \cite{sundararajan2017axiomatic} strongest overall. The mean probability-drop gap for IG is $0.2286\pm0.0125$, with corresponding mean AUC and accuracy gaps of $0.1008\pm0.0063$ and $0.2527\pm0.0154$, respectively. SmoothGrad \cite{smilkov2017smoothgrad} and input gradients also show clear positive gaps, but smaller than IG.

\begin{table}[t]
\centering
\caption{Interpretability masking effectiveness over 5 seeds. Random masking uses matched mask sizes.}
\label{tab:interp-effect}
\begin{tabular}{lccc}
\toprule
Method & Targeted drop & Random drop & Gap (targeted$-$random) \\
\midrule
Input gradients & $0.1951 \pm 0.0182$ & $0.0305 \pm 0.0011$ & $0.1645 \pm 0.0178$ \\
Integrated gradients & $0.2591 \pm 0.0130$ & $0.0305 \pm 0.0011$ & $0.2286 \pm 0.0125$ \\
SmoothGrad & $0.2021 \pm 0.0117$ & $0.0305 \pm 0.0011$ & $0.1716 \pm 0.0118$ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{JetClass-to-AspenOpenJets Shift Diagnostics}
We next evaluated each of the five seeded JetClass models on AspenOpenJets (AOJ) in an unlabeled setting, using $10^6$ AOJ jets per seed. Since AOJ is unlabeled in this setup, we report distribution-shift diagnostics and pseudo-accuracy transfer estimates derived from the JetClass corruption calibration mapping.

Table~\ref{tab:aspen-shift} shows large and consistent cross-domain shift across seeds. In particular, \texttt{prob\_l1\_drift}, \texttt{top1\_flip\_rate}, and \texttt{class\_js\_divergence} are all substantially elevated relative to in-domain clean reference behavior. The mapped pseudo-$\Delta$Acc estimates are also large; however, as discussed below, these should be treated as qualitative sensitivity indicators rather than exact transferred accuracies.

\begin{table}[t]
\centering
\caption{AspenOpenJets shift diagnostics and pseudo-accuracy transfer estimates over 5 seeds ($10^6$ AOJ jets/seed).}
\label{tab:aspen-shift}
\begin{tabular}{lccc}
\toprule
Metric & Mean $\pm$ Std & Min & Max \\
\midrule
\texttt{prob\_l1\_drift} & $0.9839 \pm 0.1026$ & $0.8878$ & $1.1494$ \\
\texttt{top1\_flip\_rate} & $0.7123 \pm 0.0298$ & $0.6656$ & $0.7374$ \\
\texttt{class\_js\_divergence} & $0.1871 \pm 0.0322$ & $0.1404$ & $0.2222$ \\
\texttt{confidence\_drop} & $0.2351 \pm 0.0514$ & $0.1481$ & $0.2803$ \\
\texttt{entropy\_shift} & $0.6920 \pm 0.1470$ & $0.4870$ & $0.8421$ \\
Pseudo $\Delta$Acc (all-metric ensemble) & $0.6325 \pm 0.0688$ & $0.5396$ & $0.7041$ \\
Pseudo expected Acc (all-metric ensemble) & $0.1135 \pm 0.0635$ & $0.0474$ & $0.1973$ \\
Pseudo $\Delta$Acc (strong 3 metrics only) & $0.4871 \pm 0.0411$ & $0.4403$ & $0.5375$ \\
Pseudo expected Acc (strong 3 metrics only) & $0.2588 \pm 0.0453$ & $0.1994$ & $0.3148$ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Sanity Checks for Aspen Pipeline}
To test whether these large AOJ shifts could be an artifact of implementation, we ran a dedicated 5-seed sanity suite covering replay consistency and output integrity. Table~\ref{tab:aspen-sanity} summarizes the outcomes.

\begin{table}[t]
\centering
\caption{Core pipeline sanity-check outcomes for Aspen evaluation across 5 seeds.}
\label{tab:aspen-sanity}
\begin{tabular}{lc}
\toprule
Check & Pass count \\
\midrule
Clean replay (accuracy consistency) & $5/5$ \\
Clean replay (AUC consistency) & $5/5$ \\
Corruption replay consistency & $5/5$ \\
Aspen output integrity (finite, normalized, size) & $5/5$ \\
Strong-metric mapping $R^2$ threshold & $5/5$ \\
Strong-metric mapping intercept sanity & $5/5$ \\
\bottomrule
\end{tabular}
\end{table}

These checks support that the pipeline is functioning as intended: clean and corruption replays match exactly across seeds, and AOJ inference outputs are numerically well-formed. Additional calibration diagnostics indicate transfer uncertainty under strong extrapolation (18 severe and 7 warning out-of-range cases over 25 seed-metric pairs, with clipping applied to an average of 1.4 metrics per seed). Therefore, we interpret mapped pseudo-accuracy as a sensitivity indicator under shift, while treating the unlabeled shift metrics themselves as the primary robust result.

\subsection{Interpretation and Implications}
These results support four conclusions for the restart baseline:
\begin{enumerate}
\item The clean offline baseline is strong and reproducible across seeds.
\item Unlabeled shift metrics based on prediction-distribution change (especially \texttt{prob\_l1\_drift}) reliably track degradation under controlled corruptions.
\item Attribution-guided perturbations are consistently more damaging than random masking at matched perturbation budgets, indicating that the attribution signals capture model-relevant constituents.
\item On AspenOpenJets, the models exhibit consistently large domain shift across seeds; sanity checks indicate this is unlikely to be a loading/checkpoint artifact, while pseudo-accuracy transfer estimates remain calibration-limited under extrapolation.
\end{enumerate}

This establishes a stable empirical baseline for the next stage of the project, where we will test whether these diagnostics can be used to drive improved training or adaptation strategies.
